---
title: "Wholesale Customer Spending Clustering"
author: "Rex Manglicmot"
date: "2022-12-19"
output: 
  github_document: 
    toc: yes

---

## Status: Continuing Working Document

## Introduction

1) FRESH: annual spending (m.u.) on fresh products (Continuous);
2) MILK: annual spending (m.u.) on milk products (Continuous);
3) GROCERY: annual spending (m.u.)on grocery products (Continuous);
4) FROZEN: annual spending (m.u.)on frozen products (Continuous)
5) DETERGENTS_PAPER: annual spending (m.u.) on detergents and paper products (Continuous)
6) DELICATESSEN: annual spending (m.u.)on and delicatessen products (Continuous);
7) CHANNEL: customers’ Channel - Horeca (Hotel/Restaurant/Café) or Retail channel (Nominal)
8) REGION: customers’ Region – Lisnon, Oporto or Other (Nominal)
Descriptive Statistics:

(Minimum, Maximum, Mean, Std. Deviation)
FRESH ( 3, 112151, 12000.30, 12647.329)
MILK (55, 73498, 5796.27, 7380.377)
GROCERY (3, 92780, 7951.28, 9503.163)
FROZEN (25, 60869, 3071.93, 4854.673)
DETERGENTS_PAPER (3, 40827, 2881.49, 4767.854)
DELICATESSEN (3, 47943, 1524.87, 2820.106)

## Loading the Libraries
```{r, message=FALSE, warning=FALSE}
#load libraries
library(tidyverse)
library(janitor)
library(scales)
library(ggplot2)

```



## Loading the Data
```{r}
#load csv file from the UCI website and store it into an object
data_orig <- read.csv('Wholesale customers data.csv')

#view first few rows of the data
head(data_orig)
```
## Cleaning the Data
```{r}
#make a copy
data <- data_orig

#make the columns to lowercase
data <- janitor::clean_names(data)

#get rid of the categorical variables since clustering deals with continuous variables
data <- data[,-c(1,2)]

#change column names to something shorter
colnames(data) <- c("fres", "milk", "groc", "fro", "depa", "deli")
```

```{r}
#check for NAs
sum(is.na(data))

#check for blanks
sum(data =='')

# one more check for blanks
which(data =='', arr.ind=TRUE)

#one last check
summary(data)
```
Great we see there are no characters or letters in our dataset since having one would throw off the summary function. 

## Exploratory Data Analysis

How much do people spend on fresh versus frozen? Let's display it. 
```{r}
#fresh and frozen scatterplot
ggplot(data, aes(fres, fro)) +
  geom_point(position = 'jitter', alpha=0.5, color= 'blue') +
  scale_y_continuous(trans = 'log10', labels = dollar) +
  scale_x_continuous(trans = 'log10', labels = dollar) +
  stat_ellipse(level = .99, color = "green") +
  theme_classic()+
labs(title = 'Wholesale Consumer Spending on Fresh Food vs. Frozen Food',
     x = 'Fresh Food Spending',
     y = 'Frozen Food Spending') +
   theme(plot.title = element_text(hjust = 0.5, size=10),
        plot.caption = element_text(hjust = 0.5))
```
```{r}
#detergeents and delicacies
ggplot(data, aes(depa, deli)) +
  geom_point(position = 'jitter', alpha=.5, color = 'red') +
  scale_y_continuous(trans = 'log10', labels = dollar) +
  scale_x_continuous(trans = 'log10', labels = dollar) +
  stat_ellipse(level = .99, color = "green") +   
  theme_classic() +
labs(title = 'Wholesale Consumer Spending on Detergents vs. Delicacies',
     x = 'Detergent Spending',
     y = 'Delicacy Spending') +
   theme(plot.title = element_text(hjust = 0.5, size=10),
        plot.caption = element_text(hjust = 0.5))
```



## Clustering



## Limitations



## Conclusions




## Inspiration for this project





